---
title: "Clustering"
author: "Erik Kahnke"
date: "August 7, 2015"
output: html_document
---

##Clustering and PCA

For this exercise I decided to use K Means clustering and PCA algorithms. For the K Means clustering, I set K=2 to assess how well K Means clustering determines reds versus whites and again where K=7 to determine if it would cluster around the seven different quality scores that were assigned to the data. When determining clusters to show the different quality scores, I chose 7 because the quality values only range from 3 to 9 in the dataset even though quality is assessed on a scale of 1 to 10. I also tried to determine if K Means clustering when K=2 is capable of sorting the higher quality wines from the lower quality wines.

```{r, include=FALSE}
wine = read.csv('wine.csv', header=TRUE)
x = wine[,-(12:13)]
x = scale(x, center=TRUE, scale=TRUE)
clust2 = kmeans(x, 2, nstart=500)
clust7 = kmeans(x, 7, nstart=500)
pc1 = prcomp(x, scale=TRUE)
loadings = pc1$rotation
scores = pc1$x
```


Below, we can see the K-Means Clustering algorithm's attempt to determine wine color. I plotted color against alcohol content to give the y-axis a continuous numerical scale.

```{r, echo=FALSE}
qplot(color, alcohol, data=wine, color=factor(clust2$cluster))
```

And now for PCA.

```{r, echo=FALSE}
qplot(scores[,1], scores[,2], data=wine, color=wine$color, xlab='Component 1', ylab='Component 2')
```

Based on the above visuals, K Means Clustering is more easily able to distinguish reds versus whites. We can see that K Means better determines wine color because there are fewer errors or incorrect classifications of red versus white.

Now lets consider how well PCA and K Means clustering determine wine by quality score. The first plot shows K Means (when K=2) attempt to sort the higher quality wines from the lower quality wines. The second plot below shows K Means (when K=7) attempt to cluster around wine quality scores and the third plot shows PCA's attempt to classify wine by quality. 

```{r, echo=FALSE}
qplot(quality, alcohol, data=wine, color=factor(clust2$cluster))
qplot(quality, alcohol, data=wine, color=factor(clust7$cluster))
qplot(scores[,1], scores[,2], data=wine, color=wine$quality, xlab='Component 1', ylab='Component 2')
```

Based on the above three visuals, we cannot say that either K Means or PCA is capable of determining wines by quality/quality score.
